Flowchart of Current Algorithm

### 2009-04-23 docs/flowchart/flowchart.tex
for spherical, we need to compute psi at time n and n+1/2
more flowchart progress
rewrote the "Changes since" section so it's accurate

%-----------------------------------------------------------------------------
% Changes Between Paper 3 and Paper 4
%-----------------------------------------------------------------------------

\section{Changes Between Paper 3 and Paper 4}

There are several changes to the algorithm since it was last detailed
in paper III.  The major changes are:

\item We have added thermal diffusion with {\tt use\_thermal\_diffusion = TRUE}.

\item We discuss spherical geometries, in paricular:
\begin{itemize}
\item We are still working out the fill, average, and interpolate stencils.
\item We are working on {\bf Advect Base} and {\bf Correct Base} details.
\item We are working on $\etarho$ details.
\item We compute the full state temperature on cell-centers after the 
advection-diffusion step using $T=T(\rho,X_k,p_0)$ by setting 
{\tt use\_tfromp = TRUE} rather than using $T=T(\rho,h,X_k)$.
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\end{itemize}

\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.

\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.

\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.
\begin{itemize}
\item First, we call {\bf Advect Base Density}, which computes a time-centered
$\etarho$.
\item Second, we call {\bf Advect Base Pressure}, which takes $\etarho$ as an
input and computes a time-centered $\psi$.
\item Third, we call {\bf Advect Base Enthalpy}, which takes $\psi$ as an input.
\end{itemize}

\item We added the volume-discrepancy term to the velocity constraint equation,
controlled by the input parameter, $f$.

\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.

\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.

\item We have given the advection scheme a face-lift, with some nifty options like
corner coupling and various PPM limiters.

\item We now update the density and enthalpy as full states with full-state fluxes, 
not as perturbational quantities.

\item We update temperature at the end of {\bf React State}.

%-----------------------------------------------------------------------------
% Changes Since Paper 4
%-----------------------------------------------------------------------------

\section{Changes Since Paper 4}

\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs immediately before {\bf Advect Base Pressure}.  Since the full
enthalpy advance ``feels'' the effects of the time-centered $\etarho$, we should
rely on a $\rho_0^{n+1}$ that also feels the effects.

\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate states, 
$(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.

\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.

@@

\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
\rightarrow [p_0^{\outp},\psi^{\outp}]$:

@@

Next, we compute $\psi^{\outp}$.  Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{equation}
\psi_j^{\outp} = 
\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip .
\end{equation}
\end{description}

\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$\\
is the process by which we update the base state enthalpy through $\dt$ in time.
\begin{description}

\item[plane-parallel:] We discretize
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to adjust the base state enthalpy
to account for large-scale mixing, reactions, and thermal diffusion.

\begin{description}
\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
\end{description}

@@

\item Define
\begin{eqnarray}
\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right] \enskip , \\
 {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
\end{eqnarray}
Here $\beta_0$ is computed as described in paper~III, Appendix C.

>>>>

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item {\tt fill\_3d} and {\tt average} algorithms.
\item We update temperature after each {\bf React State} as well as the advection step.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
\end{enumerate}
Changes which modify enthalpy update.  These are separate because in this paper,
enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
mention these items in paper IV, they should me noted in paper 4.5.
\begin{enumerate}
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}
\section{Changes Between Papers 4 and Papers 4.5}
\begin{enumerate}
\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.
\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
is in the main flow chart.
\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.
\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.
\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs after the full state density update and  before {\bf Advect Base Pressure}.  
Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
we should rely on a $\rho_0^{n+1}$ that also feels the effects.
\end{enumerate}
Changes specific to spherical:
\begin{enumerate}
\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
\item $\etarho$ and $\psi$ equations.
\end{enumerate}
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Use PPM limiters in advection.
\end{enumerate}
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume-discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\end{enumerate}

@@

\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
\rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:

@@

Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{equation}
\psi_j^{\outp,n} = 
\left(\gammabar^{\inp}p_0^{\inp}\right)_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{equation}
\psi_j^{\outp,n+\myhalf} = 
\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip .
\end{equation}
\end{description}

\item {\bf Advect Base Enthalpy}
is the process by which we update the base state enthalpy through $\dt$ in time.
The input arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
[(\rho h)_0^\outp]$

We discretize the base state enthalpy equation,
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.\\

\item {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^\inp,w_0^\inp,\psi^{\inp,n},
\psi^{\inp,n+\myhalf}] \rightarrow [(\rho h)_0^\outp]$

The base state enthalpy update now includes the area factors in the divergences:
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left \{ \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j+\myhalf} -  
          \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j-\myhalf} \right\}
+\Delta t\psi^{\inp,n+\myhalf} \enskip. 
\end{equation}
The other difference is that the interface states are computed using 
$\psi^{\inp,n}$ as the forcing.

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to adjust the base state enthalpy
to account for large-scale mixing, reactions, and thermal diffusion.
Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
\begin{description}
\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
\end{description}

@@

\item Define
\begin{eqnarray}
\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right] \enskip , \\
 {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
\end{eqnarray}
Here $\beta_0$ is computed as described in paper III, Appendix C.
